---
title: "A02a_G22630214"
author: "Christian Conroy"
date: "February 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
```

Load in the household data:

```{r, message = FALSE}
household <- read_csv("households.csv")
```

A.a -
Convert categorical variables to factors:

```{r, results='hide'}
str(household)
household$Household <- as.factor(household$Household)
household$Location <- as.factor(household$Location)
household$Household <- as.factor(household$Ownership)
```

Categorical: Household, Location, Ownership
Continuous: Family Size, First Income, Second Income, Monthly Payment, Utilities, Debt

A.b -
Nominal: Household, Location, Ownership

A.c -
Create a histogram for each of Debt: 

```{r}
hist(household$Debt)
```

A.d  - Find the maximum and minimum debt levels for the households in this sample.
```{r}
summary(household$Debt)
```
Minimum debt is $227 and maximum debt is $9104. 

A.e - Report the indebtedness levels at each of the 25th, 50th, and 75th percentiles.
```{r}
quantile(household$Debt, c(.25, .50, .70))
```

A.f - Report and interpret the interquartile range for the indebtedness levels of households
```{r}
IQR(household$Debt)
```

The difference between the upper and lower quartiles (25% and 75%) for household indebtedness is $2727
