---
title: "A02c_G22630214"
author: "Christian Conroy"
date: "February 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

C.a 

```{r}
rnorm2 <- function(n,mean,sd) { mean+sd*scale(rnorm(n)) }
set.seed(1239)
r1 <- rnorm2(100,25,4)
r2 <- rnorm2(50,10,3)
samplingframe <- c(r1,r2)
hist(samplingframe, breaks=20,col = "pink")
```
While the visual disribution appears to show right skewness, as the CLT shows, as our sample size increases, our distribution becomes approximately normal. 

C.b -
Draw 50 samples of size 15 from the sampling frame in part a, and plot the sampling
distribution of means as a histogram. 

```{r}
bsam <- replicate(50,sample(samplingframe, 15, replace = TRUE))
bsam <- as.data.frame(bsam)
bsammeans <- as.vector(sapply(bsam, mean))
hist(bsammeans, breaks=20,col = "pink")
```

C .c- Draw 50 samples of size 45 from the sampling frame in part a, and plot the sampling distribution of means as a histogram. 
```{r}
csam <- replicate(50,sample(samplingframe, 45, replace = TRUE))
csam <- as.data.frame(csam)
csammeans <- as.vector(sapply(csam, mean))
hist(csammeans, breaks=20,col = "pink")
```

C.d  - All purchase made by married female homeowners in the state of California.

```{r}
par(mfrow=c(1,2))
hist(bsammeans, breaks=20,col = "pink")
hist(csammeans, breaks=20,col = "pink")
```
The sampling plot from part a as compared to those in parts b and cappears to be approaching the most "normal"  distribution, which makes sense given the higher sample size.

C.e - Explain CLT in your own words in one or two sentences.

The CLT denotes the principle that distributions become approximately normal and sample size increases, that is to say, that the mean of all samples taken from the same population approach the approximate mean of the population. 

C.f - Does this exercise help you understand CLT? If so why? If not, why not? Restrict your response to one or two sentences.

Sure. It shows the "process" of reaching the approximately normal distribution, which is made clearer by seeing the distribution become approximately more normal as we are taking more and more samples from a population. 
