---
title: "Chapter 12 Lab Answer Key"
#author: Michael Bailey
date: 'Last Update: November 2018'
output:
  word_document: default
  pdf_document: default
---
## Preparation
```{r setup, include=TRUE, message = FALSE, warning = FALSE}
require(knitr)
library(ISLR)
library(ggplot2)
# install.packages('usmap')
library(usmap)
library(data.table)

opts_chunk$set(echo = TRUE)
options(digits = 6)
```

```{r message = FALSE, include = FALSE}
# Load in the data
data('USArrests')
```

## US Arrests: Exploratory Analysis

```{r message = FALSE}
head(USArrests)
```

```{r tidy = FALSE}

## Summary Stats
summary(USArrests)

max(USArrests$Murder)

max(USArrests$UrbanPop)

max(USArrests$Rape)

max(USArrests$Assault)


## Dimensions
dim(USArrests)

## Missingness
sapply(USArrests, function(x) sum(is.na(x)))

# Are variables related to each other
cor(USArrests)

```

Obviously the most recorded incident is the least serious one with the mean for assault at 170.8

Georgia (where I am from) is the most dangerous for murders 
91% of peiple in California are in urban
Nevada is most for rape
North Carolina is most for assault

There are 50*4 = 200 dimensions in the dataset

No missing data

Not much of a correlation between urban and crime reporting as one would expect (Rape is the closest at 0.411)

If a place has a lot of one type of crime, it makes sense that they would have a lot of another type of crime 

Clearly high between assault and murder and slightly high between rape and those two as well, but not as high as one might expect 

#### Graph Analysis 

```{r urbanrape, echo = FALSE}

##### Draw graphs to visualize correlations

# Urban Pop - Rape

ggplot(USArrests, aes(UrbanPop, Rape)) + geom_point() +
  geom_smooth()+
  labs(
    x = "Proportion living in Urban Areas",
    y= "Incidents of rape per 100,000 people"
  ) +
  scale_x_continuous(breaks = seq(20,100, by = 20)) +
  scale_y_continuous(limits = c(0,50), breaks = seq(0,50, by = 10))
```


```{r urbanmur, echo = FALSE}

# Urban Pop - Murder

ggplot(USArrests, aes(UrbanPop, Murder)) + geom_point() +
  geom_smooth()+
  labs(
    x = "Proportion living in Urban Areas",
    y= "Murders per 100,000 people"
  ) +
  scale_x_continuous(breaks = seq(20,100, by = 20)) +
  scale_y_continuous(limits = c(0,25), breaks = seq(0,25, by = 5))

```

```{r urbassault, echo = FALSE}
# Urban Pop - Assault

ggplot(USArrests, aes(UrbanPop, Assault)) + geom_point() +
  geom_smooth(method = lm)+
  labs(
    x = "Proportion living in Urban Areas",
    y= "Incidents of rape per 100,000 people"
  ) +
  scale_x_continuous(breaks = seq(20,100, by = 20)) +
  scale_y_continuous(limits = c(0,400), breaks = seq(0,400, by = 100))

```

```{r assaultrape, echo = FALSE}
# Assault - Rape

ggplot(USArrests, aes(Assault, Rape)) + geom_point() +
  geom_smooth(method = lm)+
  labs(
    x = "Assaults per 100,000 people",
    y= "Incidents of rape per 100,000 people"
  ) +
  scale_x_continuous(breaks = seq(0,400, by = 100)) +
  scale_y_continuous(limits = c(0,50), breaks = seq(0,50, by = 10))

# While the correlation is still not that high, it would appear that there may be more of a story with urban and rape. 

```


```{r AssaultMurder, echo = FALSE}
# Assault - Murder

ggplot(USArrests, aes(Assault, Murder)) + geom_point() +
  geom_smooth()+
  labs(
    x = "Assaults per 100,000 people",
    y= "Murders per 100,000 people"
  ) +
  scale_x_continuous(breaks = seq(0,400, by = 100))
scale_y_continuous(breaks = seq(20,100, by = 20))

# High corr obviously with both loess and lm showing this 
```


```{r rapemurder, echo = FALSE}
# Rape - Murder

ggplot(USArrests, aes(Rape, Murder)) + geom_point() +
  geom_smooth()+
  labs(
    x= "Incidents of rape per 100,000 people",
    y = "Murders per 100,000 people"
  ) +
  scale_x_continuous(breaks = seq(0,50, by = 10)) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5))

# High corr, but not that strong. LM would appear to show positive relationship. 
```

```{r murder, echo = FALSE}
##### Mapping the rates in the states 
# Set the row names to be the first column for use in mapping
USArrests <- setDT(USArrests, keep.rownames = TRUE)[]
names(USArrests)[1] <- "state"

# Murder
plot_usmap(data = USArrests, values = "Murder", lines = "black") + 
  scale_fill_continuous(name = "Murders per 100,000 People", label = scales::comma, low = "white", high = "red") + labs(title = "Murder Rates in the US by State") + theme(legend.position = "right")
```

```{r rape, echo = FALSE}
# Rape
plot_usmap(data = USArrests, values = "Rape", lines = "black") + 
  scale_fill_continuous(name = "Rapes per 100,000 People", label = scales::comma, low = "white", high = "red") + labs(title = "Rape Rates in the US by State") + theme(legend.position = "right")

```

```{r assault, echo = FALSE}
# Assault
plot_usmap(data = USArrests, values = "Assault", lines = "black") + 
  scale_fill_continuous(name = "Assaults per 100,000 People", label = scales::comma, low = "white", high = "red") + labs(title = "Assault Rates in the US by State") + theme(legend.position = "right")

```

```{r urbanpop, echo = FALSE}
# Urban Population
plot_usmap(data = USArrests, values = "UrbanPop", lines = "black") + 
  scale_fill_continuous(name = "% Population Urban", label = scales::comma, low = "white", high = "red") + labs(title = "% Urban by US State") + theme(legend.position = "right")

```