---
title: "Leaflet-GeoJSON"
output: html_document
---

# GeoJSON is a format for encoding a variety of geographic data structures.A GeoJSON object may represent a geometry, a feature, or a collection of features.GeoJSON supports the following geometry types: Point, LineString, Polygon, MultiPoint, MultiLineString, MultiPolygon, and GeometryCollection. Features in GeoJSON contain a geometry object and additional properties, and a feature collection represents a list of features.


```{r}
if(!require(jsonlite)){
    install.packages("jsonlite")
    library(jsonlite)
}
```

```{r}
# install.packages('devtools')
library(devtools)

if(!require(leaflet)){
    install.packages("leaflet")
    library(leaflet)
}

if(!require(shinythemes)){
    install.packages("shinythemes")
    library(shinythemes)
}
```


```{r}
library(jsonlite)
library(leaflet)
```

```{r}
# Load data from a JSON file
url <- "https://raw.githubusercontent.com/glynnbird/usstatesgeojson/master/california.geojson"
geojson <- jsonlite::fromJSON(url)

# Create a leaflet map
leaflet() %>% 
  
  # Add map tiles
  addTiles() %>%
  
  # Set the map view by longitude and latitude
  setView(lng = -98.583, lat = 39.833, zoom = 3) %>% 
  
  # Add the now-styled GeoJSON object to the map (Layer)
  addGeoJSON(geojson)
```

