---
title: "C04_G22630214"
author: "Christian Conroy"
date: "February 7, 2018"
output: html_document
---

FIX BY ADDING IN STEPWISE REGRESSION FROM HIS SLIDES AND MARKDOWNS!!!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
if (!require(dplyr)) {
  install.packages("dplyr")
  require(dplyr)
}

if (!require(ggplot2)) {
  install.packages("ggplot2")
  require(ggplot2)
}
```

```{r, include = FALSE}
setwd("/Users/chris/Documents/GeorgetownMPPMSFS/McCourtMPP/Semester4MPP/ProgrammingStats") 
benergy <- read.csv("buildingenergy.csv", header=T)

setwd("/Users/chris/Documents/GeorgetownMPPMSFS/McCourtMPP/Semester4MPP/ProgrammingStats") 
kicks <- read.csv("kicksfootball.csv", header=T)
```

```{r, include = FALSE}
colnames(kicks) <- c("Yards", "Goal", "Practiceormatch")
kicks$Goal <- ifelse(toupper(kicks$Goal)=="Y", 1,0)
```

```{r, include = FALSE}
regHL5 <- lm(HL ~ GA + WA, data = benergy)
summary(regHL5)
```

```{r, include = FALSE}
regCL5 <- lm(CL ~ GA + WA, data = benergy)
summary(regCL5)
```

```{r, include = FALSE}
practice <- kicks %>%
  filter(Practiceormatch == "P")

match <- kicks %>%
  filter(Practiceormatch == "M")
```

```{r, include = FALSE}
practlog <- glm(Goal ~ Yards, 
             data=practice,
             family=binomial)
summary(practlog)
```

```{r, include = FALSE}
matchlog <- glm(Goal ~ Yards, 
                data=match,
                family=binomial)
summary(matchlog)
```

plot HL against the most important predictor from your model in Part A
```{r}
p <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = 3, color='red')
print(p)
```


faceted by HL and RC
```{r}
q <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = .5, color='red') + facet_wrap(GD~OR)
q
```

## Fit the linear models 
HL against the most important predictor from your model in Part A
```{r}
h <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = 3, color='red') + stat_smooth(method = "lm", size = 1)
print(h)
```

faceted by HL and RC
```{r}
j <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = .5, color='red') + stat_smooth(method = "lm", size = 1) + facet_grid(GD~OR)
print(j)
```

## Second Order Polynomial
HL against the most important predictor from your model in Part A
```{r}
z <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = 3, color='red') + stat_smooth(method = "lm", formula = y ~ poly(x, 2), size = 1)
print(z)
```

faceted by HL and RC
```{r}
v <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = .5, color='red') + stat_smooth(method = "lm", formula = y ~ poly(x, 2), size = 1) + facet_grid(GD~OR)
print(v)
```

## LOESS
HL against the most important predictor from your model in Part A
```{r, warning=FALSE, message=FALSE}
k <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = 3, color='red') + stat_smooth(method = "loess", size = 1)
print(k)
```

faceted by HL and RC
```{r, warning=FALSE, message=FALSE}
m <- ggplot(benergy) + aes(x = WA, y = HL) + geom_point(size = .5, color='red') + stat_smooth(method = "loess", size = .5) + facet_grid(GD~OR)
print(m)
```

# Plot of logistic regressions
```{r}
plot(Goal~Yards,data=practice,
     xlab="Yards",
     ylab="Probability of Goal")
curve(predict(practlog,
              data.frame(Yards=x),
              type="resp"),
      add=TRUE)
par(new=TRUE)
plot(Goal~Yards,data=match,
     xlab="Yards",
     ylab="Probability of Goal")
curve(predict(matchlog,
              data.frame(Yards=x),
              type="resp"),
      add=TRUE)
```