---
title: "A03a_G22630214"
author: "Christian Conroy"
date: "February 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
```

## Part A

Download data
```{r}
temp <- tempfile(fileext = ".csv")
download.file(url = paste0("https://www2.census.gov/programs-surveys/popest/tables/2010-2015/state/totals/", "nst-est2015-01.csv"), destfile = temp, mode = "wb")
```

Read in data with appropriate rows
```{r}
all_content <- readLines(temp[1])
skip_second <- all_content[c(4, 10:60)]
statepops <- read.csv(textConnection(skip_second), header = TRUE, stringsAsFactors = FALSE)
```

Change column/variable names
```{r}
colnames(statepops) = c("state", "census", "estbase", "est2010", "est2011", "est2012", "est2013", "est2014", "est2015")
```

Fix Comma Numbers Issue
```{r}
commade <- function(numbers) {
  as.numeric(gsub(",","", numbers))
}
statepops[2:9] <- lapply(statepops[2:9], commade)
```

Save the File
```{r}
write.csv(statepops, "clean.csv", row.names = FALSE)
```

## Part B

1. Read areas.csv into a data frame. Retain only two columns: state and Land sq miles. Delete all other columns
```{r}
areas <- read.csv("areas.csv", header=T)
areas <- subset(areas, select=c("State", "Land.sq.miles"))
colnames(areas) <- c("state", "land.sq.miles")
```

2. Read clean.csv into another data frame
```{r}
clean <- read.csv("clean.csv", header = T)
```

3. Join the two data frames (or tibbles, your choice) into a third data frame. The choice of the correct join function is important. Name this joined1.
```{r}
joined1 <- left_join(clean, areas, by = "state")
head(joined1)
```

4. Create a new column in joined1 using the mutate command that computes population density (population per sq mile) based on the estimated 2015 population
```{r}
joined1 <- mutate(joined1, popdens = est2015/land.sq.miles )
```

5. Read blueredstates.csv into another data frame. Join it to joined1 and call it joined2.
```{r, warning= F}
blueredstates <- read.csv("blueredstates.csv", header=T)
names(blueredstates) <- tolower(names(blueredstates))
joined2 <- left_join(joined1, blueredstates, by = "state")
```

6. Using "piping" with %>% take joined2 as the input and pipe it into the group by operation based on the variable "Overall" (that tells you whether a state is a red or a blue state) and provide summary averages for the following variables: popden, MedianHouseholdIncome, and Land sq miles
```{r}
joined2 %>% 
  na.omit() %>%
  group_by(overall) %>% 
  summarise(mean(popdens), mean(medianhouseholdincome), mean(land.sq.miles))
```

## Part C 

1. Read clean.csv into another data frame
```{r}
clean2 <- read.csv("clean.csv", header = T)
```

2. Retain only the following variables: "est2010", "est2011", "est2012", "est2013", "est2014", and "est2015".
```{r}
clean2 <- subset(clean2, select = -c(state, census, estbase))
```

3. Provide a boxplot for the average population across all states by year. Year is on the x-axis and the average population is on the Y axis - as shown below. No labels required, default labels accepted.
```{r}
boxplot(clean2)
```

4. Make the following graph
```{r}
colnames(clean2) <- c("2010", "2011", "2012", "2013", "2014", "2015")
means <- lapply(clean2, mean)
years <- as.numeric(colnames(clean2))
plot(years, means, panel.first = grid(), col = 'red', pch = 21, bg = 'red', main = 'Mean of populations across states over the years', xlab = 'Year', ylab = 'Population', ylim=c(6000000, 6500000))
```
